\cursedforest\ - A Random Forest Implementation for ``Big'' and ``Wide'' Data \\ Supplementary Information

### 2017-02-01 doc/publications/plos16/plos16.tex
Updated part ii

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Section 2: Feature Selection on real data 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\cursedforest\ selects previously demonstrated bone-density variants}
In this section we use \cursedforest\ to identify genomic locations that are associated with a specific trait. To do this, we perform feature selection over XX genomic variants to 
identify the locations associated with Bone Mineral Density (BMD)~\cite{Duncan.2011}. We build a
model on the BMD dataset (see datasets) using a binary high or low BMD value as the label.  From this model, we output a
list of the top 10,000 contributing variants, ranked by their Gini impurity. The variant with the highest Gini impurity,
and hence importance, is ranked 1, and the variant with the lowest Gini impurity is ranked 10,000.

%DENIS: number
To test whether the associated features have biological meaning, we investigate whether the known BMD genes investigated by Duncan {\it et al.}~\cite{Duncan.2011} are enriched amongst the 
highly ranked genes. We therefore label all variants that fall within XXbp of these 32 BMD genes and observe that they are
ranked significantly higher than other locations in the genome (Mann-Whitney U, $p=1.3\times10^{-07}$), supporting our hypothesis that biologically relevant features are contributing more information to our classification model.

We therefore investigate the highest-ranked features not amongst the associated gene list by Duncan {\it et al.}~\cite{Duncan.2011}. 
%DENIS is this correct about the transcription factor binding site?
COLEC10 (collectin 10) contains variants more recently associated with BDM~\cite{Kemp2014,Liu2008}, with a verified transcription factor binding site for the known BMD gene FOXL1.
%However, we also observe that from the set of highest-ranked features, for example, the top ten, not all of them exist
%within close proximity to known BMD genes. We investigate the literature to determine the plausibility of these variants
%contributing to BMD.
%One gene is COLEC10 (collectin 10). Featuring a transcription factor binding site for FOXL1, COLEC10 is also reported as
%containing variants associated with BMD \cite{Kemp2014,Liu2008}. However there currently exists no information on how this gene
%may contribute to BMD. 
Also highly-ranked are variants in chromosome 22q11. Stagi {\it et al.}~\cite{Stagi.2010}, observe that individuals with a
microdeletion in this region have a significant reduction in bone mass. In this same region, we also observe
variants present in PRODH (proline dehydrogenase). PRODH catalyzes the first step in proline degradation, and as
proposed by Phang {\it et al.}~\cite{Phang.2008}, proline is made available by collagen degradation.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Section 2: Feature Selection on real data 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\cursedforest\ selects previously demonstrated bone-density variants}
In this section we use \cursedforest\ to identify genomic locations that are associated with a specific trait. To do
this, we perform feature selection over 7,225,008 genomic variants to identify the locations associated with Bone
Mineral Density (BMD)~\cite{Duncan.2011}. We build a model on the BMD dataset (see datasets) using a binary high or low
BMD value as the label.  From this model, we output a list of the top 10,000 contributing variants, ranked by their Gini
impurity. The variant with the highest Gini impurity, and hence importance, is ranked 1, and the variant with the lowest
Gini impurity is ranked 10,000.

To test whether the associated features have biological meaning, we investigate whether the known BMD genes investigated
by Duncan {\it et al.}~\cite{Duncan.2011} are enriched amongst the highly ranked genes. We therefore label all variants
that fall within 100 kb of these 32 BMD genes and observe that they are ranked significantly higher than other locations
in the genome (Mann-Whitney U, $p=1.3\times10^{-07}$), supporting our hypothesis that biologically relevant features are
contributing more information to our classification model.

We therefore investigate the highest-ranked features not amongst the associated gene list by Duncan
{\it et al.}~\cite{Duncan.2011}. COLEC10 (collectin 10) contains variants more recently associated with
BDM~\cite{Kemp2014,Liu2008}, with a verified transcription factor binding site for the known BMD gene FOXL1~\cite{DECODE}.
%However, we also observe that from the set of highest-ranked features, for example, the top ten, not all of them exist
%within close proximity to known BMD genes. We investigate the literature to determine the plausibility of these variants
%contributing to BMD.
%One gene is COLEC10 (collectin 10). Featuring a transcription factor binding site for FOXL1, COLEC10 is also reported as
%containing variants associated with BMD \cite{Kemp2014,Liu2008}. However there currently exists no information on how this gene
%may contribute to BMD. 
Also highly-ranked are variants in chromosome 22q11. Stagi {\it et al.}~\cite{Stagi.2010}, observe that individuals with a
microdeletion in this region have a significant reduction in bone mass. In this same region, we also observe
variants present in PRODH (proline dehydrogenase). PRODH catalyzes the first step in proline degradation, and as
proposed by Phang {\it et al.}~\cite{Phang.2008}, proline is made available by collagen degradation.
